---
title: "ECON 3880: Assignment 3"
author: "Samuel Morency - 101065894"

date: '`r Sys.Date()`'
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    df_print: paged
    number_sections: false
    theme: darkly
    highlight: breezedark
---

>Download and read the data set **SYRflights2021.rds** from Brightspace and 
save it in the same folder as your R Notebook file. Then read it into RStudio 
and call it **syr** for simplicity.

```{r}
syr <- readRDS("./SYRflights2021.rds")
```


>This data set contains flights data from the Syracuse Airport in 2021, which is 
similar to the New York flights dataset that we have seen in class and has the 
same variables. Please use an R Notebook to answer the following questions and 
submit the HTML file.

# Question 1:

>Load tidyverse and write a code to print the first 3 rows of the dataset.

```{r}
library(tidyverse)
syr %>% head(3)
```
>What are the names of columns in this dataset?

```{r}
syr %>% colnames()

# alternative:
# paste(names(syr),collapse = ", ")
```
The names of the columnns in this dataset are: year, month, day, dep_time, 
sched_dep_time, dep_delay, arr_time, sched_arr_time, arr_delay, carrier, flight,
tailnum, origin, dest, air_time, distance, hour, minute, and time_hour.

# Question 2:

>Use one of the functions we have learned so far to get the structure of the syr 
dataset and understand the variables included the dataset. What types of data 
we have in this dataset?

```{r}
syr %>% str()
```

This dataset consists of integers, characters (strings), and numeric data.

# Question 3:

>Find all flights with more than 5 hours delayed arrival that flew to New York 
City (there are 3 airports with the codes JFK, LGA, and EWR in this city).

```{r}
syr %>% filter(arr_delay > 5*60 & dest %in% c("JFK","LGA","EWR"))
```
>How many flights match these two conditions?

9 flights match these two conditions.

# Question 4:

>In the first three months of 2021, how many flights to CLT have been operated 
by American Airlines (AA) or Republic Airways (YX) with the more than 110 and 
less than 120 minutes air time? Show these cases in a table.

```{r}
syr %>% filter(year == 2021 & month <= 3 & dest == "CLT" & carrier %in% c("AA","YX") & air_time > 110 & air_time < 120)
```

23 flights.

# Question 5:

>Use 2 different approaches that we have learned in class, to create 2 new 
variables called distance_km1 and distance_km2 and convert the distance from 
miles to KM (note that 1 mile = 1.61 KM). Make sure that both variables are 
added to the syr dataset.

```{r}
syr <- syr %>% mutate(distance_km1 = 1.61 * distance)

syr$distance_km2 <- (1.61 * syr$distance)

syr
```

>Next, create a new dataset called df_new by selecting the following columns: 
MONTH, carrier, distance_km1, distance_km2, and all the columns that their name 
ends with “time”.

```{r}
df_new <- syr %>% select(month, carrier, distance_km1, distance_km2, ends_with("time"))
df_new
```

>Finally show the last 4 rows of this dataset.

```{r}
df_new %>% tail(4)
```

# Question 6:

>Using the dataset df_new, create an interactive count plot to show the number 
of observations for each carrier at different months and color them “gold”. 
Hint, make sure that the values on axes makes sense and appear correctly. In 
addition, add “Number of observations in different months for different 
carriers”, “Month of the year” and “Airline Code” as the main title of the plot 
and axes titles, respectively.

```{r}
library(plotly)

factorCols <- c("month","carrier")
df_new[factorCols] <- lapply(df_new[factorCols], as.factor)

plot1 <- df_new %>% ggplot(aes(month, carrier)) +
  geom_count(color = "gold") +
  labs(title = "Number of observations in different months for different carriers",
       x = "Month of the year",
       y = "Airline Code") 

plot1 %>% ggplotly()
```

>Overall, which carrier operated year-round and had the highest flights in all 
months?

```{r}
df_new2 <- df_new %>% 
  group_by(month, carrier) %>% 
  dplyr::summarize(count = n()) %>% 
  pivot_wider(names_from = month, 
              values_from = count, 
              names_prefix = "month_")

df_new2 <- df_new2 %>% 
  mutate(annual = rowSums(select(df_new2, starts_with("month_")), na.rm = TRUE)) %>% 
  arrange(desc(annual))

df_new2 %>% na.omit()
```
Carrier 9E operated year-round and had the highest flights in all months.

>Which carrier operated only in November and December 2021 and what was 
the total number of flights by this carrier?

```{r}
df_new2 %>% filter(if_all(month_1:month_10, is.na))
```
Carrier WN operated only in November and December 2021 and the total number of 
flights by this carrier was 144.
