---
title: "ECON 3880: Assignment 1"
author: "Samuel Morency - 101065894"

date: '`r Sys.Date()`'
output: 
  html_notebook: 
    df_print: paged
    number_sections: false
    theme: darkly
    highlight: breezedark
---

# {.tabset}

## Question 1

> Create a histogram chart for the variable **charges** and color them **blue** 
and fill in **red**.

```{r}
library(tidyverse)
library(ggplot2)

df <- read.csv("insurance.csv", header = TRUE)

factorCols <- c("sex","children","smoker","region")
df[factorCols] <- lapply(df[factorCols], as.factor)

ggplot(df, aes(charges)) + 
  geom_histogram(color = "blue",
                 fill = "red")
```
## Question 2

>Repeat the above plot again but instead of a histogram, create density plots 
and fill them according to the variable **Smoker** and also use the argument 
**alpha = 0.5** to make the densities transparent.

```{r}
ggplot(df, aes(charges, fill = smoker)) +
  geom_density(color = "blue",
               alpha = 0.5)
```
>Is there any differences between smokers and non-smokers?

The charges for smokers are significantly higher compared to non-smokers, at 
least 2-4 times larger.

## Question 3

>Is there any regional differences in density curves of **charges** among 
**smokers** and **nonsmokers**? To show these differences create a graph and 
explain if you see any differences.

```{r}
ggplot(df, aes(charges, fill = smoker)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~region)
```
While the positions on the x-axis appear to be largely the same, the height of 
the density peaks on the y-axis appear to be different. Most interestingly, out 
of the two peaks for smokers, the left peak is higher in the north, while the 
right-most peak is higher in the south.

## Question 4

>Install a package called **ggthemes** and load it into R. Next, create a 
boxplot chart for **Charges** and fill them based on **Smoker** and show the 
regional differences. Then, flip the axes and also change the theme to 
economist, which is included in the **ggtheme** package.

```{r}
library(ggthemes)

ggplot(df, aes(charges, smoker, fill = smoker)) +
  geom_boxplot() +
  facet_wrap(~region) +
  coord_flip() +
  theme_economist(dkpanel = TRUE) +
  scale_fill_economist()
```
## Question 5

>Create a **boxplot** chart to examine the link between **sex** and **charges** 
and fill them according to **smoker**.

```{r}
ggplot(df, aes(sex, charges, fill = smoker)) +
  geom_boxplot() +
  theme_economist() +
  scale_fill_economist()
```
>Next, use the same variables and create a **violin** plot instead.

```{r}
ggplot(df, aes(sex, charges, fill = smoker)) +
  geom_violin() +
  theme_economist() +
  scale_fill_economist()
```
>What does this graph show? What can be seen using this graph that is not 
observable in the boxplot?

This graph conveys similar information as the box plot, except it allows you
to see the density distribution. The box plot only shows you summary statistics.