---
title: "ECON 3880: Assignment 5"
author: "Samuel Morency - 101065894"
date: '`r Sys.Date()`'
output: 
  html_notebook: 
    toc: no
    df_print: paged
    number_sections: false
    theme: darkly
    highlight: breezedark
---

# Questions: {.tabset}

## Question 1

> Using the **statcanR** package, search for the data table that contains 
“unemployment”, “census”, and “2022” in its title, download that table and 
call it **LF**.

```{r, warning=FALSE, message=FALSE}
invisible(lapply(c("statcanR", "tidyverse", "readxl", "regclass"), 
       require, 
       character.only = TRUE))
statcan_search(c("unemployment", "census", "2022"), "eng")
```

```{r, warning=FALSE, message=FALSE}
LF <- statcan_download_data("14-10-0106-01", "eng")
LF
```
<br>

-   <div> > Clean the names of the columns</div>

```{r}
LF <- LF %>% janitor::clean_names()
LF
```
<br>

-   <div> > keep only ref_date, geo, labour_force_characteristics, 
population_centre_and_rural_areas, sex, age_group, and value</div>

```{r}
LF <- LF %>% select("ref_date", 
                    "geo", 
                    "labour_force_characteristics", 
                    "population_centre_and_rural_areas", 
                    "sex", 
                    "age_group", 
                    "value")
LF
```
<br>

-   <div> > only keep 2020 from the ref_date; “Total, all population centres 
and rural areas” from the population_centre_and_rural_areas; unemployment rate 
from labour_force_characteristics; and age groups 15-24 years and 25-44 years 
from the age_group.</div>

```{r}
LF$ref_date <- LF$ref_date %>% format("%Y")
LF <- LF %>% filter(ref_date == 2020, 
              population_centre_and_rural_areas == "Total, all population centres and rural areas",
              labour_force_characteristics == "Unemployment rate",
              age_group %in% c("15 to 24 years", "25 to 44 years"))
LF
```
<br>

-   <div> > What are the unemployment rates of males and females aged 25-44 
years old in Canada and provinces? To answer, create a table that shows the geo 
in rows, sex in the columns, and the unemployment rates in different cells</div>

```{r, warning=FALSE, message=FALSE}
uerates <- LF %>% 
  filter(sex != "Both sexes", age_group == "25 to 44 years") %>% 
  group_by(geo, sex) %>% 
  summarise(unemployment_rate = mean(value)) %>% 
  pivot_wider(names_from = sex, values_from = unemployment_rate)
uerates
```
<br>

-   <div> > Create a point plot to visualize the above table.</div>

```{r}
plot1 <- uerates %>% 
  ggplot(aes(Females, geo)) +
  geom_segment(aes(xend = Males,
                   yend = geo, 
                   linetype = 5),
               linewidth = 1) +
  scale_linetype_identity() +
  geom_point(aes(color = "Females"),
               size = 4,
             shape = 15) +
  geom_point(aes(Males, 
                 color = "Males"),
             size = 4,
             shape = 15)

plot1 <- plot1 + 
  geom_text(aes(label = abs(uerates$Females - uerates$Males) %>% round(1)),
            position = position_nudge(-(uerates$Females - uerates$Males) / 2, 0.4),
            size = 3) +
  labs(title = 'Gender gaps in Unemployment Rates', 
       subtitle = "by Geography, 25-44 year olds, Canada",
       x = 'Unemployment Rate (%)', 
       y = 'Geography', 
       colour = 'Sex')

plot1
```
<br>

-   <div> > Where do we observe the biggest gap between the females and males 
unemployment rate?</div>

```{r}
gap <- uerates %>% mutate(gap = abs(Females - Males))
gap[which.max(gap$gap),]
```
We observe the biggest gap between the female and male unemployment rates in 
`r gap[[which.max(gap$gap), "geo"]]`, at `r gap[[which.max(gap$gap), "gap"]]`%.

## Question 2

> Write code to import the excel data file named **assignment5_data** from the 
Brightspace and call it df. 

```{r}
df <- read_excel("Assignment5_data.xlsx", skip = 1)
df
```

> Estimate the following linear model $y = \beta_0 + \beta_1 * x1+\beta_2 * x2 
+ \beta_3 * x3$ and show the estimated coefficients.

```{r}
model1 <- lm(y ~ x1 + x2 + x3, data = df)
coef(model1)
```

> Which one of the independent variables are statistically insignificant?

```{r}
summary(model1)
```
A variable is statistically significant if its p-value is less than 0.05. 
Therefore, the independent variables x1 and x3 are statistically insignificant, 
while the intercept and x2 are statistically significant.

## Question 3

> Use the df data:

-   <div> > create a new variable called **y_group**, based on the values for y 
and define y_group = 1 when y > average(y) and 0 otherwise</div>

```{r}
df_q3 <- df %>% mutate(y_group = if_else(y > mean(y), 1, 0))
```

-   <div> > estimate a logistic regression using the y_group as the dependent 
variable and x1, x2, and x3 as the predictors</div>

```{r}
df_q3$y_group <- as.factor(df_q3$y_group)

model2 <- glm(y_group ~ x1 +x2 +x3, data = df_q3, family = binomial())
summary(model2)
```

-   <div> > calculate and show the confusion matrix and the accuracy rate for 
this model.</div>

```{r}
cm <- confusion_matrix(model2)
cm
```

```{r}
tn <- cm[1,1]
tp <- cm[2,2]
total <- cm[3,3]

accuracy <- (tn + tp) / total
accuracy
```
The accuracy rate for this model is ~60.7%.
