---
title: "ECON 3880: Assignment 4"
author: "Samuel Morency - 101065894"
date: '`r Sys.Date()`'
output: 
  html_notebook: 
    toc: no
    df_print: paged
    number_sections: false
    theme: darkly
    highlight: breezedark
---

> Download and read the data set **flights2022.rds** from Brightspace and save 
it in the same folder as your R Notebook file. Then read it into RStudio and 
call it **df22** for simplicity.

```{r}
df22 <- readRDS("./flights2022.rds")

factorCols <- c("year","month", "day", "carrier", "origin", "dest")
df22[factorCols] <- lapply(df22[factorCols], as.factor)
```

> Notice that this dataset is different from what we have used in the 
Assignment 3! This data set contains flights data in 2022, which is similar to 
the New York flights dataset and has been extracted using a package called 
**anyflights** in R.

# Questions: {.tabset}

## Question 1

> Using df22:

-   <div> > calculate the number and percentage of cancelled flights in each 
month (note that for cancelled flights the air_time is missing) and show them 
in a table in the R notebook </div>
```{r}
library(tidyverse)
months <- df22 %>% select(month, air_time) %>% 
  group_by(month)

summary <- months %>% summarize(flights = n())

months_na <- months %>% filter(is.na(air_time))

summary_na <- months_na %>% summarize(canceled = n())

summary <- summary %>% left_join(summary_na, by = "month") %>% 
  mutate(percent_canceled = (canceled / flights) * 100)
summary
```

-   <div> > create a bar chart to show the percentages of canceled flights in 
different months, fill them based on month </div>
```{r}
plot1 <- summary %>% ggplot(aes(month, fill = month, weight = percent_canceled)) +
  geom_bar() +
  scale_fill_hue(direction = 1)
plot1
```

-   <div> > remove the legend from the plot </div>
```{r}
plot1 <- plot1 + theme(legend.position = "none")
plot1
```

-   <div> > add "Proportion of Cancelled Flights Each Month" and "proportion 
cancelled" to the plot in appropriate places. </div>
```{r}
plot1 <- plot1 +
  labs(title = "Proportion of Cancelled Flights Each Month", 
       y = "proportion cancelled")
plot1
```

## Question 2

> Use the df22 dataset and

-   <div> > keep only the flights from JFK to LAX and IAD </div>

```{r}
JFK_to_LAX_IAD <- df22 %>% filter(origin == "JFK" & dest %in% c("LAX", "IAD"))
JFK_to_LAX_IAD
```

-   <div> > select year, origin, dest, and all the columns with a name that 
starts with **arr** </div>

```{r}
JFK_to_LAX_IAD <- JFK_to_LAX_IAD %>% select(year, origin, dest, starts_with("arr"))
JFK_to_LAX_IAD
```

-   <div> > save the results as **my_filtered_data.csv** (no need to send the 
saved data file, I only want to see the code) </div>

```{r}
write_csv(JFK_to_LAX_IAD, "my_filtered_data.csv")
```

-   <div> > what is the dimension of the saved file (write the code to get that 
information). </div>

This is how I would do it, because I know the saved file would have the same 
dimensions as the dataframe I passed into write_csv():
```{r}
dim(JFK_to_LAX_IAD)
```

But since this question explicitly asks to write code to get the dimension of 
the saved file, this is an alternative that reads the csv file and checks the 
dimensions of the result:

```{r}
read_csv("my_filtered_data.csv") %>% dim()
```
The saved file has 11243 rows, and 5 columns.

## Question 3

> How many different airports are listed as destination in df22? Write a code 
to show their names.

```{r}
dest_airports <- unique(df22$dest)
dest_airports

# alternative:
# paste(names(dest_airports),collapse = ", ")
```
```{r}
length(dest_airports)
```

There are 118 unique airports listed as destinations in df22.

## Question 4

> Write a code that shows the **number** of flights from different airports to 
MYR, OAK, SNA, and PDX airports in df22.

```{r}
to_MYR_OAK_SNA_PDX <- df22 %>% 
  filter(dest %in% c("MYR", "OAK", "SNA", "PDX")) %>% 
  group_by(origin, dest) %>% 
  summarise(number_of_flights = n()) %>% 
  pivot_wider(names_from = dest, values_from = number_of_flights, names_prefix = "flights_to_")
to_MYR_OAK_SNA_PDX
```


## Question 5

> How many missing observations do we have in each column of the df22 (note 
that we need the number of missing in every single column!)? Write a code to 
count the missing observations and show them in the notebook.

```{r}
df22_missing_obs <- df22 %>% 
  map_df(~sum(is.na(.))) %>% 
  pivot_longer(cols = everything(), 
               names_to = "column", 
               values_to = "missing_observations") %>% 
  arrange(desc(missing_observations))
# I might be able to use a '.' instead of 'everything()' here.
df22_missing_obs
```

## Question 6

> Use the df22 data:

-   <div> > drop all the flights to EGE, </div>

```{r}
df22_Q6 <- df22 %>% filter(dest != "EGE")
df22_Q6
```

-   <div> > keep only origin, dest, and distance columns, </div>

```{r}
df22_Q6 <- df22_Q6 %>% select(origin, dest, distance)
df22_Q6
```

-   <div> > exclude all the rows with any missing observation, </div>

```{r}
df22_Q6 <- df22_Q6 %>% na.omit()
df22_Q6
```
If I do it this way, no rows are excluded. This is because only the dep_time, 
dep_delay, arr_time, arr_delay, tailnum, and air_time columns contain NA values 
when a flight was canceled.

Here is an alternative solution that filters all flights that were canceled, 
assuming that is the meaning of "missing observations":

```{r}
df22_Q6 <- df22 %>% 
  filter(dest != "EGE") %>% 
  na.omit() %>% 
  select(origin, dest, distance)
df22_Q6
```


-   <div> > remove all the duplicated rows (i.e., rows with exactly the same 
values), </div>

```{r}
df22_Q6 <- df22_Q6 %>% distinct()
df22_Q6
```

-   <div> > create a wide dataset that shows the destinations in rows and the 
origins in the columns, </div>

```{r}
df22_Q6_wide <- df22_Q6 %>% pivot_wider(names_from = origin, values_from = distance, names_prefix = "origin_")
df22_Q6_wide
```

-   <div> > show the top 10 rows of the wide dataset. </div>

Here are the first 10 rows:
```{r}
df22_Q6_wide %>% head(10)
```
"Top rows" could also be interpreted to mean the highest in value, so here is 
an alternative solution showing the top 10 highest rows in terms of total 
flights:

```{r}
df22_Q6_wide %>% 
  mutate(total = rowSums(select(., starts_with("origin_")), na.rm = TRUE)) %>% 
  arrange(desc(total)) %>% 
  head(10)
```

## Question 7

> Read a data set called **Money** from the package **Ecdat**, which shows 
Money, GDP and Interest Rate in Canada and assign it to "can_money":

```{r}
can_money <- Ecdat::Money
```

-   <div> > use as.data.frame() function to convert that into a data 
frame </div>

```{r}
can_money <- can_money %>% as.data.frame()
can_money
```

-   <div> > create a new variable called my as m \* y and add that to the 
dataset </div>

```{r}
can_money <- can_money %>% 
  mutate(my = m * y)
can_money
```

-   <div> > show the first 2 rows of the data. </div>

```{r}
can_money %>% head(2)
```
