---
title: "ECON 3880: Assignment 2"
author: "Samuel Morency - 101065894"
date: '`r Sys.Date()`'
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
    df_print: paged
    number_sections: false
    theme: darkly
    highlight: breezedark
---

# [Question 1:](https://youtu.be/dQw4w9WgXcQ){target="_blank"}

> Use **sex** and **region** to create a **count plot**. Which region has the 
highest number of observations?

```{r}
library(tidyverse)
library(ggthemes)

df <- read.csv("insurance.csv", header = TRUE)

factorCols <- c("sex","children","smoker","region")
df[factorCols] <- lapply(df[factorCols], as.factor)

p1 <- ggplot(df) + 
  aes(region, sex) +
  geom_count(color = "blue")
p1
```

Southeast has the highest number of observations, 170 + 180 = `r 170 + 180`

# [Question 2:](https://youtu.be/dQw4w9WgXcQ){target="_blank"} {.tabset}

> Does the link between bmi and charges vary based on smoker?

## a)

> Create a point plot between the bmi and charges and color them based on 
smoker.

```{r}
p2 <- ggplot(df) + 
  aes(bmi, charges, color = smoker) +
  geom_point()
p2
```

## b)

> Next, add 2 linear trend lines to the plot using geom_smooth based on smoker.

```{r}
p2 <- p2 + 
  geom_smooth(method = lm)
p2
```

## c)

> Change the axes names to Body Mass Index and Dollar Charged.

```{r}
p2 <- p2 + 
  labs(x = "Body Mass Index", y = "Dollar Charged")
p2
```

## d)

> Add a title to the graph to read The link between BMI and Charges: Smokers 
and non-smokers.

```{r}
p2 <- p2 + 
  labs(title = "The link between BMI and Charges: Smokers and non-smokers.")
p2
```

## e)

> What do you conclude based on this graph?

```{r}
p2
```

From this graph I conclude that Dollars Charged has a stronger correlation with
Body Mass Index (BMI) among smokers than among non-smokers. BMI has little 
effect on Dollars Charged for non-smokers (indicated by the nearly-
horizontal trend line) whereas BMI has a significant effect on Dollars Charged 
for smokers (indicated by the steeply sloped trend line).

# [Question 3:](https://youtu.be/dQw4w9WgXcQ){target="_blank"}

> Create a stacked bar chart for each region and fill the bars based on smoker. 
Next flip the coordinates (axes) and also change the theme of the graph to 
minimal. Add the count of the observations to each bar and adjust them to see 
each number in the correct position (note that we should see 8 numbers in the 
graph!). Next, instead of letting R choose the colors for the bars, use the 
function scale_fill_brewer() and a palette called "Dark2" to change the color 
of the bars.

```{r}
p3 <- ggplot(df) + 
  aes(region, fill = smoker) +
  geom_bar(position = "stack") +
  coord_flip() +
  geom_text(stat = "count", aes(label = after_stat(count)), hjust = 1.5) +
  theme_minimal() +
  scale_fill_brewer(palette = "Dark2", direction = 1)
p3

```

# [Question 4:](https://youtu.be/dQw4w9WgXcQ){target="_blank"}

> Using the Plotly package in R, create an interactive chart that shows the 
scatter plot between bmi (in the horizontal axis), charges (in the vertical 
axis), and color the dots based on region.

```{r}
library(plotly)
p4 <- plot_ly(df, 
              x = ~ bmi, 
              y = ~ charges, 
              color = ~ region,
              type = "scatter",
              mode = "markers")
p4
```

## [Source](https://youtu.be/dQw4w9WgXcQ){target="_blank"}
